---
title: "RAM_SR_analysis"
author: "Hao Ye"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    css: styles.css
---
# Setup

Load in the packages we need and the analysis functions.

```{r setup, message = FALSE}
library(rEDM)
library(parallel)
library(quantreg)

source("my_functions.R")
```

# Pre-processing

Read the data and documentation from the RAM database website.

```{r download data, eval = FALSE}
extract_data() # pull data from RAM website
process_data() # process SR data
summarize_data()
```

Check overlap between web files and zip file from Steve.

```{r data validation, eval = FALSE}
get_doc_info()
load("stock_ids.Rdata")
both <- intersect(andi_stock_ids, web_stock_ids)
a_not_w <- setdiff(andi_stock_ids, web_stock_ids)
w_not_a <- setdiff(web_stock_ids, andi_stock_ids)
```

# Analysis

Do univariate analysis. For each stock, do simplex and s-map, find whether AR-0 or AR-1 fits better using AICc, compute simplex and s-map for 1000 AR surrogates.

```{r run analysis, eval = FALSE}
run_univariate_analysis()
append_univariate_analysis()
```

```{r collate results, warning = FALSE}
load("sr_results.Rdata")

results <- do.call(rbind, lapply(sr_results, function(x) {
    if(class(x) == "try-error")
        return(NA)
    
    # pull out EDM performance
    simplex_temp <- x$simplex_out[x$simplex_out$tp == 1,]
    smap_temp <- x$smap_out[x$smap_out$tp == 1,]
    n <- sum(is.finite(x$rec))
    best_E <- x$best_E
    simplex_rho <- simplex_temp$rho[simplex_temp$E == best_E]
    simplex_mae <- simplex_temp$mae[simplex_temp$E == best_E]
    smap_rho <- max(smap_temp$rho)
    smap_mae <- min(smap_temp$mae)
    
    # compute AR performance
    if(x$ar_0$aicc < x$ar_1$aicc)
    {
        ar_pred <- x$rec - x$ar_0$residuals
    } else {
        ar_pred <- x$rec - x$ar_1$residuals
    }
    ar_rho <- cor(x$rec, ar_pred, use = "pairwise")
    ar_mae <- mean(abs(x$rec - ar_pred), na.rm = TRUE)
    
    # compute EDM p-values
    simplex_rho_p <- (sum(simplex_rho < x$simplex_null$rho)+1) / (NROW(x$simplex_null) + 1)
    smap_rho_p <- (sum(smap_rho < x$smap_null$rho)+1) / (NROW(x$smap_null) + 1)
    simplex_mae_p <- (sum(simplex_mae > x$simplex_null$mae)+1) / (NROW(x$simplex_null) + 1)
    smap_mae_p <- (sum(smap_mae > x$smap_null$mae)+1) / (NROW(x$smap_null) + 1)
    
    return(data.frame(n = n, 
                      simplex_rho = simplex_rho, 
                      simplex_mae = simplex_mae, 
                      smap_rho = smap_rho, 
                      smap_mae = smap_mae, 
                      ar_rho = ar_rho, 
                      ar_mae = ar_mae, 
                      simplex_rho_p = simplex_rho_p, 
                      smap_rho_p = smap_rho_p))
}))
```

# Plots

## Determine how predictability changes with data length.

```{r plot simplex predictability vs. data length, fig.width = 8, echo = FALSE, results = "hide"}
par(mar = c(4, 4, 1, 1), mgp = c(2.5, 1, 0), mfrow = c(1, 2))
plot(results$n, results$simplex_rho, xlab = "Data Length", ylab = "Forecast Skill (simplex rho)")
quant_reg <- rq(simplex_rho ~ n, data = results, tau = c(0.1, 0.5, 0.9))
abline(quant_reg$coefficients[,1], lty = 2, col = "blue")
abline(quant_reg$coefficients[,2], lty = 2, col = "blue")
abline(quant_reg$coefficients[,3], lty = 2, col = "blue")

plot(results$n, results$simplex_mae, xlab = "Data Length", ylab = "Forecast Error (simplex MAE)")
quant_reg <- rq(simplex_mae ~ n, data = results, tau = c(0.1, 0.5, 0.9))
abline(quant_reg$coefficients[,1], lty = 2, col = "blue")
abline(quant_reg$coefficients[,2], lty = 2, col = "blue")
abline(quant_reg$coefficients[,3], lty = 2, col = "blue")
```

```{r plot smap predictability vs. data length, fig.width = 8, echo = FALSE, results = "hide"}
par(mar = c(4, 4, 1, 1), mgp = c(2.5, 1, 0), mfrow = c(1, 2))
plot(results$n, results$smap_rho, xlab = "Data Length", ylab = "Forecast Skill (s-map rho)")
quant_reg <- rq(smap_rho ~ n, data = results, tau = c(0.1, 0.5, 0.9))
abline(quant_reg$coefficients[,1], lty = 2, col = "blue")
abline(quant_reg$coefficients[,2], lty = 2, col = "blue")
abline(quant_reg$coefficients[,3], lty = 2, col = "blue")

plot(results$n, results$smap_mae, xlab = "Data Length", ylab = "Forecast Error (s-map MAE)")
quant_reg <- rq(smap_mae ~ n, data = results, tau = c(0.1, 0.5, 0.9))
abline(quant_reg$coefficients[,1], lty = 2, col = "blue")
abline(quant_reg$coefficients[,2], lty = 2, col = "blue")
abline(quant_reg$coefficients[,3], lty = 2, col = "blue")
```

## Comparisons of EDM performance vs. AR models

```{r plot simplex rho vs. AR rho, echo = FALSE, results = "hide"}
idx_1 <- results$n < 20
idx_2 <- (20 <= results$n) & (results$n < 30)
idx_3 <- (30 <= results$n) & (results$n < 40)
idx_4 <- 40 <= results$n

make_panel <- function(idx, plot_title = "", 
                      x_var = "ar_rho", y_var = "simplex_rho")
{
    x <- results[idx, x_var]
    y <- results[idx, y_var]
    limits <- range(c(x, y), na.rm = TRUE)
    plot(x, y, 
         xlim = limits, ylim = limits, 
         xlab = "", ylab = "", 
         main = plot_title)
    abline(a = 0, b = 1, lty = 2, col = "blue")
    return()
}

par(mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0), 
    mgp = c(2.5, 1, 0), mfrow = c(2, 2), pty = "s")

make_panel(idx_1, "Length < 20")
make_panel(idx_2, "20 <= Length 30")
make_panel(idx_3, "30 <= Length < 40")
make_panel(idx_4, "40 <= Length")

mtext("AR forecast skill (rho)", side = 1, outer = TRUE)
mtext("Simplex forecast skill (rho)", side = 2, outer = TRUE)
```

```{r plot simplex mae vs. AR mae, echo = FALSE, results = "hide"}
par(mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0), 
    mgp = c(2.5, 1, 0), mfrow = c(2, 2), pty = "s")

make_panel(idx_1, "Length < 20", x_var = "ar_mae", y_var = "simplex_mae")
make_panel(idx_2, "20 <= Length 30", x_var = "ar_mae", y_var = "simplex_mae")
make_panel(idx_3, "30 <= Length < 40", x_var = "ar_mae", y_var = "simplex_mae")
make_panel(idx_4, "40 <= Length", x_var = "ar_mae", y_var = "simplex_mae")

mtext("AR forecast error (MAE)", side = 1, outer = TRUE)
mtext("Simplex forecast error (MAE)", side = 2, outer = TRUE)
```

```{r plot s-map rho vs. AR rho, echo = FALSE, results = "hide"}
par(mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0), 
    mgp = c(2.5, 1, 0), mfrow = c(2, 2), pty = "s")

make_panel(idx_1, "Length < 20", x_var = "ar_rho", y_var = "smap_rho")
make_panel(idx_2, "20 <= Length 30", x_var = "ar_rho", y_var = "smap_rho")
make_panel(idx_3, "30 <= Length < 40", x_var = "ar_rho", y_var = "smap_rho")
make_panel(idx_4, "40 <= Length", x_var = "ar_rho", y_var = "smap_rho")

mtext("AR forecast error (rho)", side = 1, outer = TRUE)
mtext("S-map forecast error (rho)", side = 2, outer = TRUE)
```

```{r plot s-map mae vs. AR mae, echo = FALSE, results = "hide"}
par(mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0), 
    mgp = c(2.5, 1, 0), mfrow = c(2, 2), pty = "s")

make_panel(idx_1, "Length < 20", x_var = "ar_mae", y_var = "smap_mae")
make_panel(idx_2, "20 <= Length 30", x_var = "ar_mae", y_var = "smap_mae")
make_panel(idx_3, "30 <= Length < 40", x_var = "ar_mae", y_var = "smap_mae")
make_panel(idx_4, "40 <= Length", x_var = "ar_mae", y_var = "smap_mae")

mtext("AR forecast error (MAE)", side = 1, outer = TRUE)
mtext("S-map forecast error (MAE)", side = 2, outer = TRUE)
```

## P-values (probability that recruitment can be modeled using AR)

```{r plot p-values vs. data length, fig.width = 8, echo = FALSE, results = "hide"}
par(mar = c(4, 4, 1, 1), mgp = c(2.5, 1, 0), mfrow = c(1, 2))
plot(results$n, results$simplex_rho_p, xlab = "Data Length", ylab = "P-value (simplex rho)")
quant_reg <- rq(simplex_rho_p ~ n, data = results, tau = c(0.1, 0.5, 0.9))
abline(quant_reg$coefficients[,1], lty = 2, col = "blue")
abline(quant_reg$coefficients[,2], lty = 2, col = "blue")
abline(quant_reg$coefficients[,3], lty = 2, col = "blue")

plot(results$n, results$smap_rho_p, xlab = "Data Length", ylab = "Forecast Skill (smap rho)")
quant_reg <- rq(smap_rho_p ~ n, data = results, tau = c(0.1, 0.5, 0.9))
abline(quant_reg$coefficients[,1], lty = 2, col = "blue")
abline(quant_reg$coefficients[,2], lty = 2, col = "blue")
abline(quant_reg$coefficients[,3], lty = 2, col = "blue")
```

# Plots (part 2)

## Comparisons of EDM performance vs. AR models (2-, 3-, 4-step ahead forecasts)

```{r compile results 2, echo = "FALSE"}

get_results <- function(selected_tp = 2)
{
    return(do.call(rbind, lapply(sr_results, function(x) {
        if(class(x) == "try-error")
            return(NA)
        
        # pull out EDM performance
        simplex_temp <- x$simplex_out[x$simplex_out$tp == selected_tp,]
        smap_temp <- x$smap_out[x$smap_out$tp == selected_tp,]
        n <- sum(is.finite(x$rec))
        best_E <- x$best_E
        simplex_rho <- simplex_temp$rho[simplex_temp$E == best_E]
        simplex_mae <- simplex_temp$mae[simplex_temp$E == best_E]
        smap_rho <- max(smap_temp$rho)
        smap_mae <- min(smap_temp$mae)
        ar1_rho <- x$ar1_out$rho[x$ar1_out$tp == selected_tp]
        ar1_mae <- x$ar1_out$mae[x$ar1_out$tp == selected_tp]
        
        return(data.frame(n = n, 
                          simplex_rho = simplex_rho, 
                          simplex_mae = simplex_mae, 
                          smap_rho = smap_rho, 
                          smap_mae = smap_mae, 
                          ar1_rho = ar1_rho, 
                          ar1_mae = ar1_mae))
    })))
}
```

```{r plot tp 2 simplex vs. AR, fig.width = 8, echo = FALSE, results = "hide"}
results <- get_results(selected_tp = 2)

par(mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0), 
    mgp = c(2.5, 1, 0), mfrow = c(2, 2), pty = "s")

make_panel(idx_1, "Length < 20", x_var = "ar1_rho")
make_panel(idx_2, "20 <= Length 30", x_var = "ar1_rho")
make_panel(idx_3, "30 <= Length < 40", x_var = "ar1_rho")
make_panel(idx_4, "40 <= Length", x_var = "ar1_rho")

mtext("AR-1 forecast skill (rho)", side = 1, outer = TRUE)
mtext("Simplex forecast skill (rho)", side = 2, outer = TRUE)
mtext("2-step ahead forecasts", side = 3, outer = TRUE)
```

```{r plot tp 2 s-map vs. AR, fig.width = 8, echo = FALSE, results = "hide"}

par(mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0), 
    mgp = c(2.5, 1, 0), mfrow = c(2, 2), pty = "s")

make_panel(idx_1, "Length < 20", x_var = "ar1_rho", y_var = "smap_rho")
make_panel(idx_2, "20 <= Length 30", x_var = "ar1_rho", y_var = "smap_rho")
make_panel(idx_3, "30 <= Length < 40", x_var = "ar1_rho", y_var = "smap_rho")
make_panel(idx_4, "40 <= Length", x_var = "ar1_rho", y_var = "smap_rho")

mtext("AR-1 forecast skill (rho)", side = 1, outer = TRUE)
mtext("S-map forecast skill (rho)", side = 2, outer = TRUE)
mtext("2-step ahead forecasts", side = 3, outer = TRUE)
```

```{r plot tp 3 simplex vs. AR, fig.width = 8, echo = FALSE, results = "hide"}
results <- get_results(selected_tp = 3)

par(mar = c(2, 2, 1, 1), oma = c(3, 3, 3, 0), 
    mgp = c(2.5, 1, 0), mfrow = c(2, 2), pty = "s")

make_panel(idx_1, "Length < 20", x_var = "ar1_rho")
make_panel(idx_2, "20 <= Length 30", x_var = "ar1_rho")
make_panel(idx_3, "30 <= Length < 40", x_var = "ar1_rho")
make_panel(idx_4, "40 <= Length", x_var = "ar1_rho")

mtext("AR-1 forecast skill (rho)", side = 1, outer = TRUE)
mtext("Simplex forecast skill (rho)", side = 2, outer = TRUE)
mtext("3-step ahead forecasts", side = 3, outer = TRUE)
```

```{r plot tp 3 s-map vs. AR, fig.width = 8, echo = FALSE, results = "hide"}

par(mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0), 
    mgp = c(2.5, 1, 0), mfrow = c(2, 2), pty = "s")

make_panel(idx_1, "Length < 20", x_var = "ar1_rho", y_var = "smap_rho")
make_panel(idx_2, "20 <= Length 30", x_var = "ar1_rho", y_var = "smap_rho")
make_panel(idx_3, "30 <= Length < 40", x_var = "ar1_rho", y_var = "smap_rho")
make_panel(idx_4, "40 <= Length", x_var = "ar1_rho", y_var = "smap_rho")

mtext("AR-1 forecast skill (rho)", side = 1, outer = TRUE)
mtext("S-map forecast skill (rho)", side = 2, outer = TRUE)
mtext("3-step ahead forecasts", side = 3, outer = TRUE)
```

```{r plot tp 4 simplex vs. AR, fig.width = 8, echo = FALSE, results = "hide"}
results <- get_results(selected_tp = 4)

par(mar = c(2, 2, 1, 1), oma = c(3, 3, 3, 0), 
    mgp = c(2.5, 1, 0), mfrow = c(2, 2), pty = "s")

make_panel(idx_1, "Length < 20", x_var = "ar1_rho")
make_panel(idx_2, "20 <= Length 30", x_var = "ar1_rho")
make_panel(idx_3, "30 <= Length < 40", x_var = "ar1_rho")
make_panel(idx_4, "40 <= Length", x_var = "ar1_rho")

mtext("AR-1 forecast skill (rho)", side = 1, outer = TRUE)
mtext("Simplex forecast skill (rho)", side = 2, outer = TRUE)
mtext("4-step ahead forecasts", side = 3, outer = TRUE)
```

```{r plot tp 4 s-map vs. AR, fig.width = 8, echo = FALSE, results = "hide"}

par(mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0), 
    mgp = c(2.5, 1, 0), mfrow = c(2, 2), pty = "s")

make_panel(idx_1, "Length < 20", x_var = "ar1_rho", y_var = "smap_rho")
make_panel(idx_2, "20 <= Length 30", x_var = "ar1_rho", y_var = "smap_rho")
make_panel(idx_3, "30 <= Length < 40", x_var = "ar1_rho", y_var = "smap_rho")
make_panel(idx_4, "40 <= Length", x_var = "ar1_rho", y_var = "smap_rho")

mtext("AR-1 forecast skill (rho)", side = 1, outer = TRUE)
mtext("S-map forecast skill (rho)", side = 2, outer = TRUE)
mtext("4-step ahead forecasts", side = 3, outer = TRUE)
```

# Plots (part 3)
```{r compile results 3, echo = FALSE}
get_results <- function(selected_tp = 2)
{
    return(do.call(rbind, lapply(sr_results, function(x) {
        if(class(x) == "try-error")
            return(NA)
        
        # pull out EDM performance
        smap_temp <- x$smap_out[x$smap_out$tp == selected_tp,]
        n <- sum(is.finite(x$rec))
        smap_rho <- max(smap_temp$rho)
        smap_mae <- min(smap_temp$mae)
        arE_rho <- x$arE_out$rho[x$arE_out$tp == selected_tp]
        arE_mae <- x$arE_out$mae[x$arE_out$tp == selected_tp]
        
        return(data.frame(n = n, 
                          smap_rho = smap_rho, 
                          smap_mae = smap_mae, 
                          arE_rho = arE_rho, 
                          arE_mae = arE_mae))
    })))
}

```

```{r plot tp 1 s-map vs. AR-E, fig.width = 8, echo = FALSE, results = "hide"}
results <- get_results(selected_tp = 1)
idx_1 <- results$n < 20
idx_2 <- (20 <= results$n) & (results$n < 30)
idx_3 <- (30 <= results$n) & (results$n < 40)
idx_4 <- 40 <= results$n
par(mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0), 
    mgp = c(2.5, 1, 0), mfrow = c(2, 2), pty = "s")

make_panel(idx_1, "Length < 20", x_var = "arE_rho", y_var = "smap_rho")
make_panel(idx_2, "20 <= Length 30", x_var = "arE_rho", y_var = "smap_rho")
make_panel(idx_3, "30 <= Length < 40", x_var = "arE_rho", y_var = "smap_rho")
make_panel(idx_4, "40 <= Length", x_var = "arE_rho", y_var = "smap_rho")

mtext("AR-E forecast skill (rho)", side = 1, outer = TRUE)
mtext("S-map forecast skill (rho)", side = 2, outer = TRUE)
mtext("2-step ahead forecasts", side = 3, outer = TRUE)
```

```{r plot tp 2 s-map vs. AR-E, fig.width = 8, echo = FALSE, results = "hide"}
results <- get_results(selected_tp = 2)
idx_1 <- results$n < 20
idx_2 <- (20 <= results$n) & (results$n < 30)
idx_3 <- (30 <= results$n) & (results$n < 40)
idx_4 <- 40 <= results$n
par(mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0), 
    mgp = c(2.5, 1, 0), mfrow = c(2, 2), pty = "s")

make_panel(idx_1, "Length < 20", x_var = "arE_rho", y_var = "smap_rho")
make_panel(idx_2, "20 <= Length 30", x_var = "arE_rho", y_var = "smap_rho")
make_panel(idx_3, "30 <= Length < 40", x_var = "arE_rho", y_var = "smap_rho")
make_panel(idx_4, "40 <= Length", x_var = "arE_rho", y_var = "smap_rho")

mtext("AR-E forecast skill (rho)", side = 1, outer = TRUE)
mtext("S-map forecast skill (rho)", side = 2, outer = TRUE)
mtext("2-step ahead forecasts", side = 3, outer = TRUE)
```

```{r plot tp 3 s-map vs. AR-E, fig.width = 8, echo = FALSE, results = "hide"}
results <- get_results(selected_tp = 3)

par(mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0), 
    mgp = c(2.5, 1, 0), mfrow = c(2, 2), pty = "s")

make_panel(idx_1, "Length < 20", x_var = "arE_rho", y_var = "smap_rho")
make_panel(idx_2, "20 <= Length 30", x_var = "arE_rho", y_var = "smap_rho")
make_panel(idx_3, "30 <= Length < 40", x_var = "arE_rho", y_var = "smap_rho")
make_panel(idx_4, "40 <= Length", x_var = "arE_rho", y_var = "smap_rho")

mtext("AR-E forecast skill (rho)", side = 1, outer = TRUE)
mtext("S-map forecast skill (rho)", side = 2, outer = TRUE)
mtext("3-step ahead forecasts", side = 3, outer = TRUE)
```

```{r plot tp 4 s-map vs. AR-E, fig.width = 8, echo = FALSE, results = "hide"}
results <- get_results(selected_tp = 4)

par(mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0), 
    mgp = c(2.5, 1, 0), mfrow = c(2, 2), pty = "s")

make_panel(idx_1, "Length < 20", x_var = "arE_rho", y_var = "smap_rho")
make_panel(idx_2, "20 <= Length 30", x_var = "arE_rho", y_var = "smap_rho")
make_panel(idx_3, "30 <= Length < 40", x_var = "arE_rho", y_var = "smap_rho")
make_panel(idx_4, "40 <= Length", x_var = "arE_rho", y_var = "smap_rho")

mtext("AR-E forecast skill (rho)", side = 1, outer = TRUE)
mtext("S-map forecast skill (rho)", side = 2, outer = TRUE)
mtext("4-step ahead forecasts", side = 3, outer = TRUE)
```